On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2015-09-28 MS.tex
Cover letter + changes in the starred literature

\title{On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted}
\author{Anais Gibert\textasteriskcentered, Emma F. Gray, Mark Westoby, Ian J. Wright, Daniel S. Falster}
\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia\\
\textasteriskcentered Email for correspondence: \texttt{anais.gibert@gmail.com}\\
Word count: ~XXXX words}
\date{}

@@

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 274/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate (GR) has often been seen as a central element of its ecological strategy, and as ultimately linked to its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and GR, casting doubt on the capacity of traits to predict growth.

@@

\item Synthesis: We show how trait-GR correlations change in a predictable way with plant size. These  Our understanding of plant strategies should shift away from attributing species as having slow or fast GR throughout their life, in favour of attributing growth trajectories.

@@

It has previously been noted that the correlation of SLA with growth disappears among saplings and larger trees (\citealt{Wright:2010tp} for Barro Colorado Island forest, Panama), and this meta-analysis generalises that result. Several authors have suggested that the disappearance of SLA's influence on growth might be due to a disproportionate accumulation of leaf area by low SLA species \citep{Reich:1992wm,Poorter:2008iu,Wright:2010tp}, yet it remains unclear whether such a mechanism works as intended. Further, the assumed empirical relationship between SLA and total leaf area does not hold true as a generalisation across large number of species (Duursma and Falster, in review). In the model described here, the mechanism is slightly different: high-SLA species incur faster leaf turnover, the cost of which outweighs the benefits of cheap leaf construction at larger sizes.

@@

We thank Andrea Stephens for helpful comments on the meta-analyses, and Sean Gleason et al for pre-publication access to their data. Charles Warren and Jordi Martìnez-Vilalta kindly provided additional information about their published results. This work was funded by the Australian Research Council through a fellowship to MW, a discovery grant to DF and a Macquarie University Research Fellowships to EG.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig2.pdf}
\caption{The range of plant sizes and stage recoded in original publications, compared to standardised thresholds used in the analysis. Panels are for
a) age, b) height and c) diameter. Horizontal line segments indicate the size range (max., min.) for the plants used to measure growth. Study IDs correspond to the reference number in our dataset. Colours denote stage information given by the authors. The solid vertical lines indicate the thresholds used in our analysis. "*" indicated when a stage reassignment is needed. "+" on left side of the graph indicate cases where plants had different growth form. Results shown here are the ''complete'' dataset (see methods for details).}
\label{fig:fig2}
\end{figure}

>>>>

\title{On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted}
\author{Anais Gibert\textasteriskcentered, Emma F. Gray, Mark Westoby, Ian J. Wright, Daniel S. Falster}
\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia\\
\textasteriskcentered Email for correspondence: \texttt{anais.gibert@gmail.com}\\
Word count: ~5538 words (without supplementary material, Fig and ref)}
\date{}

@@

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 232/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate (GR) has often been seen as a central element of its ecological strategy, and as ultimately linked to its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and GR, casting doubt on the capacity of traits to predict growth.

@@

\item Synthesis: We show how trait-GR correlations change in a predictable way with plant size. Our understanding of plant strategies should shift away from attributing species as having slow or fast GR throughout their life, in favour of attributing growth trajectories.

@@

It has previously been noted that the correlation of SLA with growth disappears among saplings and larger trees (\citealt{Wright:2010tp} for Barro Colorado Island forest, Panama), and this meta-analysis generalises that result. Several authors have suggested that the disappearance of SLA's influence on growth might be due to a disproportionate accumulation of leaf area by low SLA species \citep{Reich:1992wm,Poorter:2008iu,Wright:2010tp}, yet it remains unclear whether such a mechanism works as intended. Further, the assumed empirical relationship between SLA and total leaf area does not hold true as a generalisation across large number of species (\citeauthor{Duursma-2015}, in review). In the model described here, the mechanism is slightly different: high-SLA species incur faster leaf turnover, the cost of which outweighs the benefits of cheap leaf construction at larger sizes.

@@

We thank Andrea Stephens for helpful comments on the meta-analyses, and Sean Gleason et al for pre-publication access to their data. Charles Warren and Jordi Martìnez-Vilalta kindly provided additional information about their published results. This work was funded by the Australian Research Council through a fellowship to M.W, a discovery grant to D.S.F and a Macquarie University Research Fellowships to E.F.G.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig2.pdf}
\caption{The range of plant sizes and stage recoded in original publications, compared to standardised thresholds used in the analysis. Panels are for
a) age, b) height and c) diameter. Horizontal line segments indicate the size range (max., min.) for the plants used to measure growth. Study IDs correspond to the reference number in our dataset. Colours denote stage information given by the authors. The solid vertical lines indicate the thresholds used in our analysis. "*" indicated when a stage reassignment is needed. "+" on left side of the graph indicate cases where plants came from different growth form. Results shown here are the ''complete'' dataset (see methods for details).}
\label{fig:fig2}
\end{figure}
